# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
First, we unzip and read in the files from 'activity.zip'.

```{r read}
unzip("./activity.zip")
activity <- read.csv("activity.csv")
```

we make the date variable a 'real' time variable recognized by R.
```{r date}
activity$Date <- strptime(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r steps}
summarize_act <- ddply(activity, .(date), summarize, Total=sum(steps), Mean=mean(steps), Median = median(steps))

hist(summarize_act$Total, breaks=20)

print("Mean: ")
mean(summarize_act$Total, na.rm=TRUE)
print("Median: ")
median(summarize_act$Total, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r activity}
act_avg <- with(activity, tapply(steps, interval, mean, na.rm=TRUE))
average <- data.frame(interval=rownames(act_avg), mean=as.numeric(act_avg))
with(average, plot(mean, interval, type="l"))
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
